The evolution of gene expression in the term placenta of viviparous mammals

### 2016-04-18 manuscript/manuscript.Rnw
discuss muglia review of P4

In both ruminants and primates, the placenta is capable of producing
estrogen and progesterone
\cite{Strauss.ea1996:Placentalsteroidhormonesynthesis}. High relative
expression of \textit{CYP19A1} (Supplemental
\cref{fig:estrogen_evolution_tree}), the enzyme that synthesizes
estrogen from androstenedione is seen in \human, \bonobo, \cow, and
\horse, though not in \sheep, \smonkey. %
%
\textit{HSD3B1}, one of two enzymes that synthesizes progesterone from
pregnalone
(\cite{Simard.ea2005:Molecularbiologyof3beta-hydroxysteroid},
Supplemental \cref{fig:progesterone_evolution_tree}) is also highly
expressed in \human, \bonobo, \cow, \horse, and \dog, but is not
highly expressed in \smonkey. The diversity of \textit{CYP19A1} and
\textit{HSD3B1} expression levels in in primates and ungulates seen
here points to the need for more complete sampling of placentas at
different timepoints and in different clades.

\DLA{discuss progesterone withdrawal and the lou muglin review}

>>>>

In both ruminants and primates, the placenta is capable of producing
estrogen and progesterone
\cite{Strauss.ea1996:Placentalsteroidhormonesynthesis}. High relative
expression of \textit{CYP19A1} (Supplemental
\cref{fig:estrogen_evolution_tree}), the enzyme that synthesizes
estrogen from androstenedione is seen in \human, \bonobo, \cow, and
\horse, though not in \sheep, \smonkey. %
%
\textit{HSD3B1}, one of two enzymes that synthesizes progesterone from
pregnalone
(\cite{Simard.ea2005:Molecularbiologyof3beta-hydroxysteroid},
Supplemental \cref{fig:progesterone_evolution_tree}) is also highly
expressed in \human, \bonobo, \cow, \horse, and \dog, but is not
highly expressed in \smonkey, \sheep, \pig, \elephant, \armadillo, or
\opossum. This is consistent with the previously observed local
progesterone withdrawal in \sheep, and the leuteo-placental shift of
progesterone expression in boreoeutheria in contrast to afrotherians
which do not show such a shift
\cite{Swaggart.ea2015:Genomicsofpretermbirth}. The diversity of
\textit{CYP19A1} and \textit{HSD3B1} expression levels in in primates
and ungulates seen here points to the need for more complete sampling
of placentas at different timepoints and in different clades.
